---
title: "CLASS8"
author: "Mengdan Liu"
date: "February 6, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
OUR TURN TO TRY K MEAN ANALYSIS
```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```

```{r}
y <- kmeans(x, centers = 2, nstart=20)
y
plot(x, col = y$cluster)
```
The next one ot learn is hierarchical clustering in R
```{r}
# First we need to calculate point (dis)similarity
# as the Euclidean distance between observations
dist_matrix <- dist(x)
# The hclust() function returns a hierarchical
# clustering model
hc <- hclust(d = dist_matrix)
# the print method is not so useful here
hc

plot(hc)

cutree(hc, h = 6)

```

Dendrograms appear to relate data points based on their proximity to a cluster. Clusters are formed based on relative euclidean distance. This is a problem related to linking methods. In other words, how the hcluster program makes groups. Taking a look at a few of them right now.

```{r}
hc.complete = hclust(d = dist_matrix, method = "complete")
hc.average = hclust(d = dist_matrix, method = "average")
hc.single = hclust(d = dist_matrix, method = "single")

plot(hc.complete)
plot(hc.average)
plot(hc.single)
```

Time for some hands on work.

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)

#QUESTION
#Identify 2-3 clusters using hclust, dist, plot, and cutree
dist_matrix2 <- dist(x)
hc2 <- hclust(d=dist_matrix2, method = "complete")
cut_hc2 <- cutree(hc2, h = 1.75)
plot(hc2)
cut_hc2
plot(x, col=cut_hc2)
```

Moving onto PCA

```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV",
 row.names=1) 
head(mydata)
```
```{r}
pca <- prcomp(t(mydata), scale=TRUE)
summary(pca)
plot(pca$x)
```

TIME FOR UNSUPERVISED WORK

```{r}
#READING DATA
foo <- read.csv("data/UK_foods.csv", row.names=1)
head(foo)
```
```{r}
#OBSERVING DIFFERENCES IN DATA
barplot(as.matrix(foo), beside = FALSE, col=rainbow(nrow(foo)))
```
```{r}
pairs(foo, col=rainbow(10), pch=16)
```

```{r}
pca <- prcomp(t(foo))
summary(pca)
```

```{r}
plot(pca$x[,1], pca$x[,2], xlab = "PC1", ylab="PC2", xlim=c(-270,500))
col.vec <- c("orange", "red", "blue", "green")
text(pca$x[,1], pca$x[,2], colnames(foo), col = col.vec)
```

