---
title: "CLASS9"
author: "Mengdan Liu"
date: "February 8, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing Wisconsin Cancer Data
```{r}
fna.data <- "data/WisconsinCancer.csv"
wisc.df <- read.csv(fna.data, header=TRUE)
```

Removing unnecessary data columns and applying rownames from data frame source.
```{r}
wisc.data <- wisc.df[,3:32]
rownames(wisc.data) <- wisc.df$id
```

Identifying number of malignant or benign cases of cancer.
```{r}
table(wisc.df$diagnosis)
```

How many features are "mean" values?
```{r}
names <- colnames(wisc.data)
length(grep("_mean", names, ignore.case = TRUE))
```
Another example where we shrink the dataset and return column name by indexing
```{r}
indexes <- grep("_mean", colnames(wisc.data), ignore.case=TRUE)
colnames(wisc.data)[indexes]
```

PCA ANALYSIS BEGINS
Checking data to see if scaling is necessary
```{r}
round(apply(wisc.data, 2, mean), 2)
print("STANDARD DEVIATIONS")
round(apply(wisc.data, 2, sd), 2)
```
Performing PCA & plotting analysis
```{r}
wisc.pr <- prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)
```

BIPLOT METHOD
```{r}
biplot(wisc.pr)
```

Standard PCA plot 1 & 2 / 1 & 3
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], xlab= "PC1", ylab = "PC2", col = wisc.df$diagnosis)
plot(wisc.pr$x[,1], wisc.pr$x[,3], xlab= "PC1", ylab = "PC3", col = wisc.df$diagnosis)
```

Variance exercises
```{r}
pr.var <- wisc.pr$sdev^2
pve <- round(pr.var/sum(pr.var)*100, 2)

plot(pve, type = "o", xlab="Principal Component", ylab = "Proportion of variance explained", ylim = c(0,100))

##Scree Barplot next
barplot(pve, ylab = "Percent of variance explained", names.arg = paste0("PC", 1:length(pve)), las = 2, axes = FALSE)
axis(2, at = pve, labels = round(pve))

```

Clustering in Principal Component space
```{r}
d <- dist (wisc.pr$x[,1:2])
hc <- hclust(d,method="ward.D2")
plot(hc)
grp3 <- cutree(hc, k=3)
table(grp3)
```
Plotting PCA with color by group as determined by hcluster analysis
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], xlab="PC1", ylab = "PC2", col = grp3)
```
Cross tabulation
```{r}
diagnosisM <- wisc.df$diagnosis == "M"
table(grp3, diagnosisM)
```

USING THE PREDICT FUNCTION
```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc

#APPLYING DATASET TO EXISTING PCA PLOT
plot(wisc.pr$x[,1:2], col = wisc.df$diagnosis)
points(npc[,1], npc[,2], col = c("green", "blue"), pch=16)
```


