---
title: "Class11&12: Structural Bionformatics (Part2) for drug discovery"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Clean up our protein target structure

First we download a target(i.e. receptor) structure from the pdb database. We will pick PDB ID "1hsg"


```{r}
#Loading bio3d tools
library(bio3d)
```

```{r}
#Getting target structure
pdb.code <- "1hsg"
file.name <- get.pdb(pdb.code)
#Reading pdb structure into pdb file.
hiv <- read.pdb(file.name)
```

Extract the protein only segment of this PDB entry and write out a new PDB format file. We will also do the same for the ligand.

```{r}
#extracting the protein section only.
protein <- trim(hiv, "protein")
ligand <- trim(hiv, "ligand")

#wrting to new pdb files.
write.pdb(protein, file="1hsg_protein.pdb")
write.pdb(ligand, file="1hsg_ligand.pdb")
```

##Convert our docking results for viewing in VMD
```{r}
res <- read.pdb("all.pdbqt", multi=TRUE)
res
```
```{r}
write.pdb(res,file = "results.pdb")
```

##Quantitiatively determining the goodness of fit for our ligand results
```{r}
res2 <- read.pdb("all.pdbqt", multi=TRUE)
ori <- read.pdb("1hsg_ligand.pdbqt")
rmsd(ori, res2)
```


