---
title: "CLASS14"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Reading in CSV data on experimental dexamethasone treatments
```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <- read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```
Examining data
```{r}
head(counts)
head(metadata)
```

How many genes are in this dataset?
```{r}
nrow(counts)
```



Extracting controls
```{r}
control <- metadata[metadata[, "dex"]== "control",]
control.mean <- rowSums(counts[,control$id])/ncol(control)
names(control.mean) <- counts$ensgene
```

Extracting treated
```{r}
treated <- metadata[metadata[,"dex"]=="treated",]
treated.mean <- rowSums(counts[,treated$id]/ncol(treated))
names(treated.mean) <- counts$ensgene
```

Combining mean counts for record keeping purposes
```{r}
meancounts <- data.frame(control.mean, treated.mean)
```
Checking sums of means of each patient group
```{r}
colSums(meancounts)
```

Creating a scatter plot of mean treated vs mean control.
```{r}
plot(meancounts$control.mean, meancounts$treated.mean)
plot(meancounts$control.mean, meancounts$treated.mean, log="xy")
```

Switching to log2fold change to observe differentially expressing genes between the two populations. 

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

Getting rid of genes with means of zero. To avoid NaN problems. Barry's way.
```{r}
#Both means have to be nonzero, iterate and get a list of indices, excluding zero means in the indice list.
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)

```

```{r}
up.ind <- mycounts$log2fc>2
down.ind <- mycounts$log2fc<2

table(up.ind)
table(down.ind)
```

##Annotations

```{r}
anno <- read.csv("annotables_grch38.csv")
head(anno)
```
Checking anno dimensions
```{r}
dim(anno)
```
Let's try out the **merge()** function to add the annotation data to our **mycounts** results.

```{r}
mycounts.anno <- merge(x=mycounts, y = anno, by.x = "row.names", by.y = "ensgene")
head(mycounts.anno)
```

Trying out BioConductor annotation packages
```{r}
source("http://bioconductor.org/biocLite.R")
biocLite("AnnotationDbi")
biocLite("org.Hs.eg.db")

```
```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```

```{r}
columns(org.Hs.eg.db)
```

```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
```


##Moving onto DESeq2
Initializing DESeq2
```{r}
library(DESeq2)
citation("DESeq2")

```

Build a DEseq dataset
```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

Run a DESeq Analysis
```{r}
dds <- DESeq(dds)
```

Getting our results
```{r}
res <- results(dds)
res
```

Analyzing results
```{r}
summary(res)
```

##Volcano plots
```{r}
plot(res$log2FoldChange, -log(res$padj))
```
Adding Color to it
```{r}
mycols <- rep("gray", nrow(res))
mycols[abs(res$log2FoldChange)>2]="red"
subset.inds <- res$padj<.01&abs(res$log2FoldChange)>2
mycols[subset.inds] <- "blue"
plot(res$log2FoldChange, -log(res$padj), col = mycols)
```

