---
title: "Class 6 HW"
author: "Mengdan Liu"
date: "January 28, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Class 6 Homework - Making a plot function

  This function has two input modes. In the first, the user must input a PDB code. The function will then output the A chain of the protein's B value plotted against its secondary structure. The second input mode again requires a PDB code, but the user can also specify which chain they'd like to plot. The function then outputs the specified chain's B value against its secondary structure.
```{r}

#Function relies on functions from bio3d library.
library(bio3d)

#plotBFactor is a function that accepts a PDB iD as argument and outputs a graph of the corresponding protein's B-factor or the amount of wiggle atoms have in a structure(?)
plotChainBFactor <- function(pdbNum, chainLetter = "A"){
  #Assigning pdb coordinate file to variable s
  inPDB <- read.pdb(pdbNum)
  inChain <- chainLetter
  #Assigns chainA atom coordinates to var s.chainA using trim.pdb
  inPDB.chain <- trim.pdb(inPDB, chain=inChain, elety="CA")
  #Assigning chainA atom b values to s.b
  inPDB.bVals <- inPDB.chain$atom$b
  #Plotting b values against chainA secondary structure using plot3b
  plotb3(inPDB.bVals, sse = inPDB.chain, typ="l", ylab="Bfactor")
  
}
plotChainBFactor("4AKE")
plotChainBFactor("4AKE", "B")
```


A small error is output after the second function call. I think this is because I'm attempting to import the same PDB file twice (to the same variable?) but this does not seem to affect the function's functionality.